Unsupervised embeddings of organizations capture the latent structure of scientific mobility

### 2020-06-10 paper/current/main.tex
revise introduction text

When scientists move, they bring their knowledge with them, and so understanding the flows of scientists allows us to understand the flow of knowledge around the world. 
However, given its complexity~\autocite{robinson2019mobility, vannoorden2012mobility, deville2014career, ackers2008mobility, scott2015dynamics}, new approaches are needed to make sense of landscape of global scientific mobility. 
A basic component of mobility is distance--people are more likely to move nearby than to cross continents.
Although many alternative ways to quantify the distance exist~\autocite{boschma2005proximity}, geographic distance remains the unambiguous and foundational quantity.
In turn, models of human mobility, such as the Gravity~\autocite{zipf1946gravity} and Radiation models ~\autocite{simini2012universal}, rely almost exclusively on geographic distance.
However, distance is not always so straightforward as simple geography. 
Geographic distance alone fails to capture the host of historical, linguistic, cultural, and cognitive distances that also drive mobility. 
For instance, Australia may be physically close to Indonesia, but in terms of shared language, history, and culture, they are essentially neighbors to the United Kingdom.
Then, how can we quantify meaningful, useful distance between institutions, cities, and countries?

Here, we propose a neural embedding model as an efficient means of encoding an effective measure of distance using large-scale data on mobility trajectories, one that encodes complex latent structure.
It has been demonstrated that the neural embedding model preserves semantic relationships between words in a dense, continuous and meaningful low-dimensional vector representation~\autocite{mikolov2013word2vec}, and has been applied to reveal insights in text~\autocite{linzhuo2020hyperbolic, tshitoyan2019mat2vec, garg2018gender, kozlowski2018geometry, hamilton2016diachronic}, documents~\autocite{le2014doc2vec, nakandala2016twitch}, and networks~\autocite{perozzi2014deepwalk, grover2016node2vec}.
In addition to encoding latent structure, this approach is computationally efficient, scalable for large datasets, robust to noise, and allows for a range of unique analyses that leverage their semantic properties using simple vector arithmetic.

As scientist's careers progress, they move between organizations or pick up additional affiliations, forming co-affiliation or \textit{mobility trajectories} that can be derived from their publication histories~\autocite{robinson2019mobility, sugimoto2017mostimpact}.
By treating geo-located organizations analogously to words and trajectories analogously to sentences, we learn vector-space embedding of organizations using a neural embedding method (Fig.~\ref{fig:methods:mobility_sentence}).
In the resulting embedding, each organization corresponds to a vector, and vectors relate to one another \textit{semantically}--just as with language, organizations are \textit{semantically} similar when they share similar co-affiliation \textit{contexts}. 
We measure the semantic distance between organization vectors using \textit{cosine distance}, and use this to model distance in the landscape of scientific mobility.

@@

We use the trajectories of the three million name-disambiguated authors who were \textit{mobile}---having more than one affiliation---between 2008 and 2019, as evidenced by their publications indexed in the Web of Science database.
We adopt the Gravity Model of mobility~\autocite{zipf1946gravity, barbosa2018human}, one of the most broadly applied mobility models~\autocite{curiel2018citygravity, lewer2008immigrationgravity, jung2008highwaygravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity} to assess how well the embedding captures mobility. After fitting the parameters of the gravity model, we use it to derive the \textit{expected flux} $\hat{T_{ij}}$ between organizations, defined as,

@@

Where the \textit{expected flux}, $\hat{T_{ij}}$, or expected number of authors co-affiliated with a pair organizations during the time period is proportional to the organization's sizes (both mobile and non-mobile authors, see Data and Methods), $m_{i}$ and $m_{j}$, and decays as a function of their distance, $f(r_{ij})$.
We consider two distances, the geographic distance between the organizations as a baseline, and the \textit{embedding distance} defined as the cosine distance between the organization's vector in the trained embedding space.
We argue that the performance of a distance measure can be observed from their ability to explain expected mobility between organizations and to make predictions using the gravity model.
For geographic distance, we define $f(r_{ij})$ as the standard power-law function, and for the embedding distance, we use the exponential function, selected as the best performing for each case (Fig.~\ref{fig:supp:predict_emb_bymodel} and Fig.~\ref{fig:supp:predict_geo_bymodel}).

>>>>

How far apart are two places?
The question is surprisingly hard to answer when it involves human mobility.
While geographic distance may have driven human movements in past eras, it holds less importance to a world connected by rapid transport, complex international relations, and global trade.
For example, a person currently living in Australia is more likely to migrate to the United Kingdom, a country with similar language and culture, than to much closer and more populous countries like Indonesia and the Philippines. 
Similarly, a prospective student living in South Korea is more likely to attend university in Canada rather than the much closer though diplomatically isolated North Korea. 
Despite its complexity, the ubiquitious models of human mobility such as the Gravity~\autocite{zipf1946gravity} and Radiation~\autocite{simini2012universal} model continue to predict mobility as a function of geographic distance.
Although alternatives exist to quantify \textit{functional} distances not based on geography~\autocite{boschma2005proximity, brown1970functional, brown1970migration, kim2018functional}, geographic distance remains the unambiguous and foundational quantity for modeling mobility.

Here, we propose a neural embedding model that uses real-world mobility trajectories to encode the complex structure of mobility into a dense, continuous, and meaningful vector-space representation, from which a \textit{functional distance} between places can be derived.
This approach, origionally designed to create embeddings of words~\autocite{mikolov2013word2vec} has been used to reveal insights in texts~\autocite{linzhuo2020hyperbolic, tshitoyan2019mat2vec, garg2018gender, kozlowski2018geometry, hamilton2016diachronic}, documents~\autocite{le2014doc2vec, nakandala2016twitch}, and networks~\autocite{perozzi2014deepwalk, grover2016node2vec}.
In addition to providing a data-driven functional distance, this approach is computationally efficient, scalable for large datasets, robust to noise, and allows for a range of unique analyses that leverage their semantic properties using simple vector arithmetic.

We use scientific mobility (scientific co-affiliation) as a case study to demonstrate the neural embedding approach. 
Scientific mobility has become increasingly globalized over time~\autocite{czaika2018globalisation}, and has been shown to drive innovation~\autocite{braunerhjelm2020labor, kaiser2018innovation, armano2017innovation}, impact~\autocite{sugimoto2017mostimpact, petersen2018multiscale, jonkers2013return, franzoni2014advantage}, collaboration~\autocite{rodrigues2016mobility}, and the diffusion of knowledge~\autocite{braunerhjelm2020labor, azoulay2011diffusion, morgan2018prestige}.
However,  given its complexity~\autocite{robinson2019mobility, vannoorden2012mobility, deville2014career, ackers2008mobility, scott2015dynamics}, new approaches are needed to make sense of landscape of global scientific mobility.

As scientist's careers progress, they move between organizations or pick up additional simultaneous affiliations forming co-affiliation or \textit{mobility trajectories}.
These mobility trajectories, represented as ordered lists of organizations, can be derived from affiliations listed on researcher's publication histories~\autocite{robinson2019mobility, sugimoto2017mostimpact} (Fig.~\ref{fig:methods:mobility_sentence}).
We use the trajectories of the three million name-disambiguated authors who were \textit{mobile}---having more than one affiliation---between 2008 and 2019, as evidenced by their publications indexed in the Web of Science database.
A vector-space embedding or organizations is learned by using mobility trajectories as input to the standard skip-gram negative sampling, or \textit{word2vec} neural-network architecture.
In the resulting embedding, each organization corresponds to a vector, and vectors relate to one another based on the likelihood of organizations appearing near one another in the provided mobility trajectories. 
We calculate the pairwise \textit{embedding distance} between organization as the cosine distance between their correponding vectors in the embedding.

@@

We assess the performance of the embedding distance relative to geographic distance by evaluating their ability to explain and predict real-world mobility as parameters in the Gravity Model of mobility~\autocite{zipf1946gravity, barbosa2018human}.
The Gravity Model, taking inspiration from Newton's law of gravity, is one of the most broadly applied mobility models~\autocite{curiel2018citygravity, lewer2008immigrationgravity, jung2008highwaygravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity} . After fitting the parameters of the gravity model, we use it to derive the \textit{expected flux} $\hat{T_{ij}}$ between organizations, defined as,

@@

Where the \textit{expected flux}, $\hat{T_{ij}}$, or expected number of authors co-affiliated with a pair organizations during the time period is proportional to the organization's sizes (both mobile and non-mobile authors, see Data and Methods), $m_{i}$ and $m_{j}$, and decays as a function of their distance, $f(r_{ij})$.
The distance function is defined in terms of either the geographic distance between organizations orthe embedding distance. 
For geographic distance, we define $f(r_{ij})$ as the standard power-law function, and for the embedding distance, we use the exponential function, selected as the best performing for each case (Fig.~\ref{fig:supp:predict_emb_bymodel} and Fig.~\ref{fig:supp:predict_geo_bymodel}).
We argue that the performance of a distance measure can be observed from its ability to explain expected mobility between organizations and to make predictions using the gravity model.
